Proof No Diamond (3 layers)

### 2023-10-23 Paper/Paper.tex
part2

Now we try to show the part 2 of No Diamond theorem. Again, we begin with 3 layers with only one output node $z$.

@@

\begin{lemma}
	A node $b_i$ in the hidden layer will not be opened in Theorem \ref{no_diamond_2} iff one of the following two happen:

@@

Because for input nodes in $I$, they also do not have compensating path pair, we can define the sign as in Definition \ref{sign_of_nodes}:

\begin{definition}[Sign of Nodes in $I$]\label{sign_of_nodes_in_I}
	We define a sign function $S$ on $I$ such that: 	(1). $S(n)=0$ if all path from $n$ to $z$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $z$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=-1$ if all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight.

	For an input node $a_i$ in $I$:
	\begin{enumerate}
		\item  $S(a_i)=0$ if 
		for all $b_j, w(a_i b_j)\cdot w(b_j c) = 0$

		\item $S(a_i)=-1$ if for all $b_j\ w(a_i b_j)\cdot w(b_j c) \leq 0$ and there exists 
		$j, w(a_i b_j)\cdot w(b_j c) < 0$ 
	\end{enumerate}
\end{definition}

\begin{lemma}
	node $z$ can get its maximal value if for all $a_i$ in $I$, $a_i$ gets its maximal if $S(a_i)=1$ or $a_i$ gets its minimal if $S(a_i)=-1$.
\end{lemma}

@@

	We use $a_S$ where $S\subseteq K$ to refer the input vector $\langle a_k\rangle_{a_k\in S}$. Then $a_I\oplus a_J = a_K$ and $c=c(a_K)=c(a_I,a_J)$.

	This is because \begin{align*}
		\max_{a_K} c(a_K) = \max_{a_J}\max_{a_I} c(a_I,a_J)
	\end{align*}

	For any fixed input of $a_J$, $c(\cdot, a_J)$ can be regarded as a DNN $D'$ with input nodes $a_i\in I$ by regarding all $a_J$ and their propagation in hidden layers as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply the No Diamond theorem, part 1 to get that: (for fixed $a_J$) $c(\cdot,a_J)$ can get its maximal value if for all $a_i$ in $I$, $a_i$ gets its maximal if $S(a_i)=1$ or $a_i$ gets its minimal if $S(a_i)=-1$.

	If we use $\max(\tilde{a}_I$ to denote above values, then we have $$\max_{a_I}c(a_I,a_J)=c(\max(\tilde{a}_I),a_J).$$ And therefore, $\max c=\max_{a_J} c(\max\tilde{a}_I,a_J)$
\end{proof}

Because for any method, like MILP, the upper bound computed will not be smaller that the actual maximal value, so we only need to show, $\mathrm{UB}_{\mathrm{MILP}}\leq \max c$ in any case.

@@

By the algorithm, $$\mathrm{UB} = B_z+\sum_{b_h\in O}w_{ab_h}\ReLU(b_h(\bar{a}_I,\bar{a}_J))+$$

>>>>

Now we try to show the part 2 of No Diamond theorem. Again, we begin with 3 layers with one output node $z$.

@@

\begin{lemma} \label{lem:open_node}
	A node $b_i$ in the hidden layer will not be opened in Theorem \ref{no_diamond_2} iff one of the following two happen:

@@

Now we need the function in Definition \ref{sign_of_nodes}. The difference is that now it is a partial function.

\begin{definition}\label{sign_of_nodes_in_I}
	We define a partial sign function all nodes such that: 	(1). $S(n)=0$ if all path from $n$ to $z$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $z$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=-1$ if all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight.

	For an input node $a_i$ in $I$:
	\begin{enumerate}
		\item  $S(a_i)=0$ if 
		for all $b_j, w(a_i b_j)\cdot w(b_j z) = 0$

		\item  $S(a_i)=1$ if for all $b_j, w(a_i b_j)\cdot w(b_j z) \geq 0$ and there exists 
		$j, w(a_i b_j)\cdot w(b_j z) > 0$

		\item $S(a_i)=-1$ if for all $b_j\ w(a_i b_j)\cdot w(b_j z) \leq 0$ and there exists 
		$j, w(a_i b_j)\cdot w(b_j z) < 0$ 
	\end{enumerate}
\end{definition}

Because for input nodes in $I$, they do not have compensating path pair, so the sign function is defined on
all nodes in $I$.

We use $a_S$ where $S\subseteq K$ to refer the input vector $\langle a_k\rangle_{a_k\in S}$. Then $a_I\oplus a_J = a_K$ and $z=z(a_K)=z(a_I,a_J)$.		Let $a_I^*$ be the input vector that for all $a_i$ in $I$, $a_i$ gets its maximal (if $S(a_i)=1$) or $a_i$ gets its minimal (if $S(a_i)=-1$).

\begin{lemma} \label{lem:reach_max}

	1. Node $z$ can get its maximal value if $a_I=a_I^*$

	2. Similarly, for every node $b$ in the hidden layer (notice that $S(b)$ is always defined since there is only one path from $b$ to $z$), for any fixed input of $a_J$,  $b(\cdot,a_J)$ will get its max/min (if $S(b)=1$ or $-1$) if $a_I=a_I^*$.
\end{lemma}

@@

	1. This is because \begin{align*}
		\max_{a_K} z(a_K) = \max_{a_J}\max_{a_I} z(a_I,a_J)
	\end{align*}

	For any fixed input of $a_J$, $z(\cdot, a_J)$ can be regarded as a DNN $D'$ with input nodes $a_i\in I$ by regarding all $a_J$ and their propagation in hidden layers as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply the No Diamond theorem, part 1 to get that: (for fixed $a_J$) $z(\cdot,a_J)$ can get its maximal value if for all $a_i$ in $I$, $a_i$ gets its maximal if $S(a_i)=1$ or $a_i$ gets its minimal if $S(a_i)=-1$.

	If we use $\max(\tilde{a}_I$ to denote above values, then we have $$\max_{a_I}z(a_I,a_J)=z(\max(\tilde{a}_I),a_J).$$ And therefore, $\max z=\max_{a_J} z(\max\tilde{a}_I,a_J)$

	2. This is simpler because for each fixed input of $a_J$, $$b= B'+\sum_{a_i\in I} w_{ba_i}a_i,$$ where $B'$ is a constant which is the sum of rest terms, including bias and $a_J$. By Lemma \ref{lemma1}, we can easily see the claim of this lemma is true.
\end{proof}

Because for any  sound method, like MILP, the upper bound computed will not be smaller that the actual maximal value, so we only need to show, $\mathrm{UB}_{\mathrm{MILP}}\leq \max c$ in any case.

@@

By the algorithm, \begin{align*}
	\mathrm{UB} = B_z+\sum_{b_h\in O}w_{zb_h}\ReLU(b_h(\bar{a}_I,\bar{a}_J))+\sum_{b_k\notin O}w_{zb_k}\mathrm{Approx}(b_k(\bar{a}_I,\bar{a}_J)).
\end{align*} Here, $\mathrm{Approx}$ is the approximation function of $\mathrm{ReLU}$. In our case (LP), it will equal to $\ReLU$ if $w_{zb_k}<0$, or a linear function if  if $w_{zb_k}>0$. $\mathrm{ReLU}$ and $\mathrm{Approx}$ are both non-decrease function.

Then by Lemma \ref{lem:reach_max}, for any fixed $a_J$, for all nodes $b$ in the hidden layer, $b$ will get its max/min (if $S(b)=\mathrm{sign}(w_{zb})=1,-1$) if $a_I=a_I^*$. Hence we have \begin{align*}
	\mathrm{UB} \leq  B_z+\sum_{b_h\in O}w_{zb_h}\ReLU(b_h(a_I^*,\bar{a}_J))+\sum_{b_k\notin O}w_{zb_k}\mathrm{Approx}(b_k(a_I^*,\bar{a}_J)).
\end{align*} Meanwhile, when $b_k$ reaches its max or min, we have $\mathrm{Approx}(b_k)=\ReLU(b_k)$, hence we have: \begin{align*}
	\mathrm{UB} \leq & B_z+\sum_{b}w_{zb}\ReLU(b(a_I^*,\bar{a}_J))\\
	= & \max_{a_I} z(a_I,\bar{a}_J)\\
	\leq & \max_{a_J}\max_{a_I} z(a_I,a_J)\\
	= & \max_{a_K} z(a_K) = \max z
\end{align*}
